---
title: "Initial Data Load"
author: "Scott Sutherland"
date: "October 24, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Import Challenge Data

Load the challenge data from the xlsx file into a variable named `mb`

```{r load, include=FALSE}
library(readxl)
library(feather)
library(dplyr)
setwd('~/src/moneyball-challenge')
mb <- tbl_df(read.csv('protected-data/MLB_Challenge_Data 2015 upd 11-2-16.csv', stringsAsFactors = FALSE, na.strings = "NA"))
```

```{r mbDimensions}
dim(mb)
```

# Perform some column interpretation

The DayNight column has either a 'D' for day or 'N' for night. We are going to convert that to an 'at_night' indicator with 1 indicating a night game.

```{r convertDayNightToIndicator}
mb <- mutate(mb, game.at_night = DayNight == "N")
```

The `HitterStatus` column is split into three indicators `game.status.available`, `game.status.dl`, and `game.status.brv`.

```{r convertHitterStatusToIndicator}
mb <- mutate(mb, game.status.available = HitterStatus == "A",
             game.status.dl = startsWith(HitterStatus, "D"),
             game.status.brv = HitterStatus == "BRV")
```

The `PlayToday_Likelihood` column is split into three indicators `game.in_lineup_high`, `game.in_lineup_low`, and `game.in_lineup_unsure`.

```{r convertPlayToday_LikelihoodToIndicator}
mb <- mutate(mb, game.in_linup.high = PlayToday_Likelihood == "HIGH",
             game.in_linup.low = PlayToday_Likelihood == "LOW",
             game.in_linup.unsure = PlayToday_Likelihood == "UNSURE")
```

The `Hitter.Pos` and `Fanduel.Pos` columns have been split into `game.hitter.<POS>` and `game.hitter.fanduel.<POS>` for all available position values as well as `game.hitter.corner_if` comprised of first and third basemen, `game.hitter.outfield` comprised of left, right, center field, and `game.hitter.infield` comprised of first, second, third, and shortstop.

```{r convertHitter.PosToIndicators}
mb <- mutate(mb, game.hitter.1b = Hitter.Pos == "1B",
             game.hitter.2b = Hitter.Pos == "2B",
             game.hitter.3b = Hitter.Pos == "3B",
             game.hitter.c = Hitter.Pos == "C",
             game.hitter.cf = Hitter.Pos == "CF",
             game.hitter.dh = Hitter.Pos == "DH",
             game.hitter.if = Hitter.Pos == "IF",
             game.hitter.lf = Hitter.Pos == "LF",
             game.hitter.of = Hitter.Pos == "OF",
             game.hitter.rf = Hitter.Pos == "RF",
             game.hitter.ss = Hitter.Pos == "SS",
             game.hitter.corner_if = Hitter.Pos == "3B" || Hitter.Pos == "1B",
             game.hitter.outfield = Hitter.Pos == "RF" || Hitter.Pos == "CF" || Hitter.Pos == "LF",
             game.hitter.infield = Hitter.Pos == "1B" || Hitter.Pos == "2B" || Hitter.Pos == "SS" || Hitter.Pos == "3B")

mb <- mutate(mb, game.hitter.fanduel.1b = Fanduel.Pos == "1B",
             game.hitter.fanduel.2b = Fanduel.Pos == "2B",
             game.hitter.fanduel.3b = Fanduel.Pos == "3B",
             game.hitter.fanduel.c = Fanduel.Pos == "C",
             game.hitter.fanduel.of = Fanduel.Pos == "OF",
             game.hitter.fanduel.ss = Fanduel.Pos == "SS")
```

The `PitcherSide` column is converter into `game.pitcher.is_lefty` if the pitcher is left hand dominant.

```{r convertPitcherSideToIndicator}
mb <- mutate(mb, game.pitcher.is_lefty = PitcherSide == "L")
```

The `Hitter.Pitcher.B.T` column is split into the nine combinations of right, left, and switch batter and pitcher combinations called `game.matchup.<hitter-side>.<pitcher-side>`. Additionally, we created `game.matchup.on_power_side` where the matchup was dominant hitter side on dominant pitcher side, `game.matchup.switch_hitter` where a switch hitter was in the matchup, and `game.matchup.switch_pitcher` where a switch pitcher was in the matchup.

```{r convertHitter.Pitcher.B.TToIndicator}
mb <- mutate(mb, game.matchup.switch.switch = Hitter.Pitcher.B.T == "BHB-BHP",
             game.matchup.switch.lefty = Hitter.Pitcher.B.T == "BHB-LHP",
             game.matchup.switch.righty = Hitter.Pitcher.B.T == "BHB-RHP",
             game.matchup.lefty.switch = Hitter.Pitcher.B.T == "LHB-BHP",
             game.matchup.lefty.lefty = Hitter.Pitcher.B.T == "LHB-LHP",
             game.matchup.lefty.righty = Hitter.Pitcher.B.T == "LHB-RHP",
             game.matchup.righty.switch = Hitter.Pitcher.B.T == "RHB-BHP",
             game.matchup.righty.lefty = Hitter.Pitcher.B.T == "RHB-LHP",
             game.matchup.righty.righty = Hitter.Pitcher.B.T == "RHB-RHP",
             game.matchup.on_power_side = Hitter.Pitcher.B.T == "RHB-RHP" || Hitter.Pitcher.B.T == "LHB-LHP",
             game.matchup.switch_hitter = startsWith(Hitter.Pitcher.B.T, "BHB"),
             game.matchup.switch_pitcher = endsWith(Hitter.Pitcher.B.T, "BHP"))
```

The season is broken down into three sections `game.week.early` for the first four weeks, `game.week.callups` for the last four weeks.

```{r}
mb <- mutate(mb, game.week.early = Week <= 4,
             game.week.callups = Week >= 24)
```


# Change the name of some columns

We update all the "general" columns to be prefixed with 'game'

```{r changeGeneralNames, include=FALSE}
mb <- rename(mb, game.week = Week,
             game.bop = BOP
             )
```


We update all the weather related data to have a weather prefix.

```{r changeWeatherNames, include=FALSE}
mb <- rename(mb, weather.humidity = humidity,
             weather.pressure = pressure,
             weather.condition = condition,
             weather.temperature = temp,
             weather.change.of.rain = chanceofrain,
             weather.out.to.left.impact = outtoleftimpact,
             weather.out.to.right.impact = outtorightimpact)
```

We update the head 2 head names to prefix with 'h2h'. e.g. The `PA.H2H` column becomes `h2h.pa`.

```{r changeH2HNames, include=FALSE}
mb <- rename(mb, h2h.pa = PA.H2H,
             h2h.hits = H.H2H,
             h2h.ab = AB.H2H,
             h2h.bavg = BAVG.H2H,
             h2h.slg = SLG.H2H,
             h2h.hr = HR.H2H,
             h2h.woba = wOBA.H2H,
             h2h.so = SO.H2H,
             h2h.well.hit.avg = WHAvg.H2H)
```



## Split test and training sets

Split the test and training sets into distinct variables

```{r split}
attach(mb)
test <- (is.na(Actual_PTS))
mb.test <- mb[test,]
nrow(mb.test)
mb.train <- mb[!test,]
nrow(mb.train)
```

# Store the test and train sets

Write the test and train sets to the protected-data directory for consumption by other processes.

```{r save}
write_feather(mb.test, 'protected-data/mb.test.feather')
write_feather(mb.train, 'protected-data/mb.train.feather')
```

We can now use the test and training sets individually.
